SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-05 design_arch.tex
gino check this

As discussed in Section~\ref{sec:Background-Perceptron}, perceptron learning
is a light-weight mechanism to pull together disparate forms of information
and synthesize a decision from them.  In this case, we leverage a variety of
different relative characteristics of a given prefetch, such as speculation
depth, page address and offset, into a binary useful/non-useful prediction.
In the remainder of this section, we discuss the the general structure of PPF
for enhancing a given underlying prefetcher.  We will show that PPF is a
generalized prefetch filtering mechanism that may be adapted to any prefetcher
with appropriate modifications.

@@

The Perceptron Filter and the steps used to performed inference 
and training are shown in figure~\ref{fig:PPF_Datapath}. The filter is organized as tables of 
perceptron weights, with a separate table of weights for each feature. 
For a design using `N' number of features, its weights will be arranged
in $N$ different tables. Each feature's value is used to index 
into the corresponding table with $K$ entries.  Since each table is
indexed independently, different numbers of bits can be used for indexing,
thus giving each table a different size.  The variable indexing can be
determined by studying the features and fine-tuning them empirically to
achieve a good accuracy versus hardware overhead trade-off.  Exact details of
our PPF implementation are given under ``Area Overheads'' in
Section~\ref{sec:Method-Overheads}.

A single entry in the table corresponds to a perceptron weight.  Each weight
is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
beginning of a program's execution. We empirically observe that 5 bits of
index achieves a good trade-off between hardware overhead and accuracy.\newline

@@

The prefetcher is triggered on every L2 cache
demand access.  At that point, the base prefetcher has the option to perform a
prefetch -- if it does, then it has a choice of how many cache lines to
prefetch.  The suggested prefetches can be either placed in the L2 cache or
the last-level cache based on the prefetcher's internal confidence mechanism.

When the base prefetcher is triggered, it begins to suggest candidates for
prefetching.  All suggested prefetches pass through the perceptron
filter to decide whether they are qualified for final prefetching.  The filter
decides whether to prefetch a candidate by looking at the microarchitectural
state, \textit{i.e.} the features, at that instant.  Each feature is hashed to
form an index into a table of up to 4,096 entries dedicated to that feature.

Each feature's confidence contributes to 
the final decision of whether to prefetch a cache line, with the summation 
of all features' confidences denoting the overall confidence for a prefetch.  
The final sum is compared against two thresholds: PERC\_THRESHOLD\_HI and  
PERC\_ THRESHOLD\_ LO to decide the placement of the data in the cache 
hierarchy. Prefetches that have a confidence above the higher threshold 
are placed into the L2 cache since the observed features strongly recommend  
the prefetch. A prefetch that results in a confidence between 
PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI are placed in the larger 
last level cache, as the filter is moderately confident of the cache line's 
use, but seeks to reduce pollution in the smaller L2 cache. Finally, the 
remaining prefetches that are below the lower threshold are not prefetched 
due to a lack of correlation between the current observed features and 
the prediction from the base prefetch engine.\newline

>>>>

As discussed in Section~\ref{sec:Background-Perceptron}, perceptron learning
is a light-weight mechanism to pull together disparate forms of information
and synthesize a decision from them.  In this case, we leverage a variety of
different relative characteristics of a given prefetch, such as speculation
depth, page address and offset, into a binary useful/non-useful prediction.
In the remainder of this section, we discuss the general structure of PPF
for enhancing a given underlying prefetcher.  We will show that PPF is a
generalized prefetch filtering mechanism that may be adapted to any prefetcher
with appropriate modifications.

@@

The Perceptron Filter structure and the steps used to performed inference 
and training are illustrated in figure~\ref{fig:PPF_Datapath}. 
The filter is organized as a set of tables, where each table entry holds a  
perceptron weight. 
%There exists a distinct table for each of the features. redundant with next sentence?

For a configuration using $N$ number of features, 
their corresponding weights will be distributed
in $N$ different tables. Each feature is used to index 
the corresponding table with $K$ entries.  Each table is
indexed independently, and the size of table varies, hence, different
number of bits are used to index different the tables.

%ETG: so not all tables have K entries? K is not a constant,

Variable indexing can be
determined by studying the features and tuning them empirically to
achieve a good tradeoff between high accuracy and hardware overhead.
A detail explanation of the storage overhead of
our PPF implementation can be found in
Section~\ref{sec:Method-Overheads}.

A single entry in the table corresponds to a perceptron weight.  Each weight
is a 5 bit counter saturating at -16 and +15 and initialized to zero at the
beginning of an applications's execution. We empirically observe having 5 bit weights,
yields a good trade-off between hardware overhead and accuracy.\newline

@@

The prefetcher is triggered on every L2 cache
demand access.  At that point, the base prefetcher has the option to perform a
prefetch -- if it does, then it makes the decision of how many cache lines to
prefetch.  The suggested prefetches can be either placed in the L2 cache or
the last-level cache based on a confidence value.

When the base prefetcher is triggered, it begins to suggest candidates for
prefetching.  All suggested prefetches pass through the perceptron
filter to decide whether they are qualified for final prefetching. %final prefetching -_-  
The filter makes the final decision to prefetch a \textit{prefetch candidate}
suggested by the base prefetcher.
Each of the feature is used to index a distinct table, 
the corresponding eeights are summed. The sum denotes the overall confidence for the prefetch. 
The value is thresholded against two diffent magnitudes:

%The filter
%decides whether to prefetch a candidate by looking at the microarchitectural
%state, \textit{i.e.} the features.  Each feature is used as an index 
%to its corresponding table, each table can have up 4,096 entries.
%Each feature's confidence contributes to 
%the final decision of whether to prefetch a cache line, with the summation 
%of all features' confidences denoting the overall confidence for a prefetch.  
%The final sum is compared against two thresholds:

PERC\_THRESHOLD\_HI and  
PERC\_ THRESHOLD\_ LO. 
%to decide the placement of the data in the cache 
%hierarchy. 
Prefetches that have a confidence above the higher magnitude 
are placed into the L2 cache since the obtained summed suggests with high
confidence this would be a useful prefetch.
A prefetchs which features result in a confidence value between  
%A prefetch that results in a confidence between 
PERC\_ THRESHOLD\_ LO and PERC\_ THRESHOLD\_ HI is allocated in the larger 
last-level cache, the filter is moderately confident of use of this 
cache line, but seeks to reduce pollution in a significantly smaller L2 cache. 
Lastly, prefetches which features yield a confident value lower than 
PERC\_ THRESHOLD\_ LO are not prefetched. \newline 
%due to a lack of correlation between the current observed features and 
%the prediction from the base prefetch engine.\newline
